---
title: "Response to Youngflesh and Lynch"
output: html_document
---

First let's load the data from Youngflesh and Lynch:

```{r, warning=FALSE, message=FALSE}
if (!file.exists("generated-data/youngflesh.rda")) {
  setwd("youngflesh-analysis")
  source("Scripts/Analysis.R")
  setwd("..")
  rm(list = ls())
}
load("generated-data/youngflesh.rda")
gpdd <- readRDS("youngflesh-analysis/Data/gomp-base-hat.rds")
```

Let's create a function that simulates a constant population growth rate with observation error and returns the maximum observed growth rate:

```{r}
library(ggplot2)
max_rmax <- function(rmax = 0.2, obs_sd = 0.3, nyears = 25, ...) {
  x <- 0 + rmax * seq(0, nyears-1) + rnorm(nyears, -0.5 * obs_sd^2, obs_sd)
  x_r <- log(exp(x[2:length(x)])/exp(x[1:(length(x) - 1)]))
  rmax_hat <- max(x_r, na.rm = TRUE) 
  data.frame(rmax_hat)
}
```

Now we will run our function over a set of parameter values:

```{r out, cache=TRUE}
pars <- expand.grid(i = seq_len(2000), obs_sd = c(0.05, 0.1, 0.2, 0.4), 
  nyears = 25, rmax = c(0.5, 1, 2))
out <- plyr::mdply(pars, max_rmax) %>% 
  mutate(rmax_facet = paste("r[max] =", rmax))
```

And make a quick plot of the simulation output with ggplot:

```{r, dependson="out", cache=TRUE}
ggplot(out, aes(as.factor(obs_sd), rmax_hat / rmax)) + 
  geom_violin(fill = "grey60", col = "grey60") +
  ggsidekick::theme_sleek() + # https://github.com/seananderson/ggsidekick
  geom_hline(aes(yintercept = 1), lty = 2) +
  ylab(expression(widehat(r[max])/r)) +
  xlab("Observation error SD (on log scale)") +
  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 10, 20, 40), limits = c(1, 5)) +
  theme(axis.title.x = element_text(size = 9)) +
  facet_grid(. ~ rmax, labeller = label_bquote(cols = r==.(rmax))) 
ggsave("obs-err.pdf", width = 5.7, height = 2.5)
obs_sim <- out
```

This next code chunk fits models with Stan to the GPDD populations that were flagged in Youngflesh and Lynch with a random walk and fixed observation error:

```{r}
library(dplyr)
library(rstan)
rstan_options(auto_write = TRUE)

if (!file.exists("stan-obs-rw.rds")) {
  stan_model_obs_rw <- stan_model(file = "rw-obs.stan") 
  saveRDS(stan_model_obs_rw, file = "stan-obs-rw.rds")
} else {
  stan_model_obs_rw <- readRDS("stan-obs-rw.rds")
}

gpdd <- readRDS("youngflesh-analysis/Data/gpdd-clean.rds")
load("generated-data/youngflesh.rda")
gpdd <- inner_join(gpdd, select(f_data, main_id, Rho, p10))

gpdd <- filter(gpdd, p10 > 0.5)
# length(unique(gpdd$main_id))

if (!file.exists("stan-obs-rw.rds")) {
  out <- lapply(unique(gpdd$main_id), function(ii) {
    x <- filter(gpdd, main_id == ii)
    m <- sampling(stan_model_obs_rw,
      data = list(N = nrow(x), y = log(x$population_untransformed), 
        sigma_obs = 0.2),
      iter = 2000, chains = 4, control = list(adapt_delta = 0.95))
    
    s <- rstan::extract(m, pars = "U")$U
    diff_m <- matrix(ncol = ncol(s) - 1, nrow = nrow(s))
    for (i in 2:ncol(s))
      diff_m[, i - 1] <- s[, i] - s[, i - 1]
    
    medians <- apply(diff_m, 2, median)
    max_id <- which(medians == max(medians))
    
    q <- as.data.frame(t(quantile(diff_m[, max_id],
      probs = c(0.025, 0.25, 0.5, 0.75, 0.975))))
    q$main_id <- ii
    list(r_max = q, U = s, max_diff_post = diff_m[, max_id])
  })
  saveRDS(out, file = "out.rds")
} else {
  out <- readRDS("out.rds")
}

rmax_obs <- plyr::ldply(out, function(x) x$r_max) %>%
  inner_join(select(f_data, main_id, Rho, p10, common_name, r_max)) %>%
  mutate(id = paste(common_name, main_id)) %>%
  mutate(dif = `50%` - Rho) %>%
  mutate(id = paste0(common_name, " / ", main_id)) %>%
  mutate(id = forcats::fct_reorder(id, dif))
```

Let's make a quick plot of the model output and count some values:

```{r}
ggplot(rmax_obs, aes(`50%`, id)) + geom_point() +
  geom_segment(aes(y = id, yend = id,
    x = `2.5%`, xend = `97.5%`)) +
  geom_point(aes(x = Rho), col = "red", pch = 3) +
  geom_point(aes(x = r_max), pch = 21) +
  ggsidekick::theme_sleek() +
  theme(panel.grid.major.y = element_line(colour = "grey90", linetype = 1)) +
  ylab("") + xlab("Maximum per-capita growth rate")
ggsave("rw-obs-maxr.pdf", width = 6, height = 5)

filter(rmax_obs, `50%` > Rho)
filter(rmax_obs, `25%` > Rho)
filter(rmax_obs, `2.5%` > Rho)

rmax_obs_p <- plyr::ldply(out, function(x) {
  p <- mean(x$max_diff_post < f_data$Rho[f_data$main_id == x$r_max$main_id[[1]]])
  data.frame(prob = p, main_id = x$r_max$main_id[[1]])
}) 
rmax_obs_p %>% arrange(prob)
filter(rmax_obs_p, prob < 0.5)

rmax_obs_p %>% arrange(prob)
filter(rmax_obs_p, prob < 0.5)

filter(rmax_obs, r_max > Rho)
```

Now let's generate a final combined figure for the paper:

```{r}
add_label <- function(label = "", xfrac = -0.02, yfrac = -0.04, pos = 4, ...) {
  u <- par("usr")
  x <- u[1] + xfrac * (u[2] - u[1])
  y <- u[4] - yfrac * (u[4] - u[3])
  text(x, y, label, pos = pos, ...)
}
g <- ggplot(obs_sim, aes(as.factor(obs_sd), rmax_hat / rmax)) + 
  geom_violin(fill = "grey60", col = "grey60") +
  facet_grid(. ~ rmax)
gg <- ggplot_build(g)
d <- gg$data[[1]]

pdf("fig1.pdf", width = 6, height = 5)
layout(rbind(c(1, 4), c(2, 4), c(3, 4)))
par(mar = c(0,0,0,0.5), oma = c(3, 2.8, 0.5, 11), tck = -0.025, mgp = c(2, 0.45, 0), 
  col.axis = "grey25", col = "grey25")
par(cex = 0.8)
par(xpd = FALSE)
ii <- 1
bean <- function(x) {
  if (ii == 1) scale_ <- 0.6
  if (ii == 2) scale_ <- 0.7
  if (ii == 3) scale_ <- 0.85
  plot(1, 1, type = "n", xlim = c(0.4, 4.2), ylim = c(0.8, 4.1), 
    xlab = "", ylab = "", yaxt = "n", xaxs = "i", yaxs = "i", las = 1,
    xaxt = "n", ann = FALSE, bty = "n")
  for (i in unique(x$x)) {
    xx <- filter(x, x == i)
    polygon(c(xx$x - xx$violinwidth*scale_, rev(xx$x + xx$violinwidth*scale_)), 
      c(xx$y, rev(xx$y)),
      col = "grey58", border = NA, lwd = 0.1)
  }
  box(col = "grey50")
  add_label(LETTERS[[ii]], xfrac = 0.01, yfrac = 0.11, cex = 1.2)
  add_label(paste0("     r = ", unique(obs_sim$rmax)[[unique(x$PANEL)]]), 
    xfrac = 0.01, yfrac = 0.114, cex = 1, col = "grey40")
  ii <<- ii + 1
  axis(2, col = "grey50", las = 1, at = c(1, 2, 3))
  abline(h = 1, col = "grey60", lty = 2, lwd = 1.2)
}
plyr::d_ply(d, "PANEL", function(x) bean(x))
par(xpd = NA)
par(mgp = c(2, 0.35, 0))
axis(1, col = "grey50", at = 1:4, labels = unique(obs_sim$obs_sd))
mtext("Observation CV", side = 1, line = 1.7, cex = 0.9)
mtext("Maximum observed r / r", side = 2, line = 1.5, outer = TRUE, cex = 0.9)
par(xpd = FALSE)
rm_ <- rmax_obs
rm_$id <- sub("Cabbage root fly or maggot", "Cabbage root fly", rm_$id)
rm_$id <- sub("Pine looper or Bordered white", "Pine looper", rm_$id)
rm_$id <- sub("Unknown", "Carrot psyllid", rm_$id)
rm_$id <- gsub("  ", " ", rm_$id)
rm_$id <- gsub(" / ", " (", rm_$id)
rm_$id <- paste0(rm_$id, ")")
rm <- arrange(rm, dif)
plot(1, 1, xlim = c(0, 7), ylim = c(1, nrow(rm_)), type = "n",
  ann = FALSE, ylab = "", xlab = "", bty = "n", xaxt = "n", yaxt = "n")
points(rm_$`50%`, 1:nrow(rm_), pch = 19, col = "grey30")
segments(rm_$`2.5%`, 1:nrow(rm_), rm_$`97.5`, 1:nrow(rm_), col = "grey30")
points(rm_$r_max, 1:nrow(rm_), col = "grey30", bg = "#FFFFFF", pch = 21)
points(rm_$Rho, 1:nrow(rm_), pch = 4, col = "red")
box(col = "grey50")
axis(2, side = 4, at = 1:nrow(rm_), labels = rm_$id, las = 1, tcl = 0, col = "grey50", cex.axis = 0.9)
axis(1, col = "grey50", at = seq(0, 8, 2))
add_label("D", xfrac = 0.82, yfrac = 0.035, cex = 1.2)
par(xpd = NA)
mtext("Per-capita pop. growth rate (r)", side = 1, line = 1.7, cex = 0.9, adj = 0.2)
text(3.6, 4.5, expression("Cole's"~rho), pos = 4, cex = 0.9)
text(2.5, 16, "Maximum r", pos = 4, cex = 0.9)
arrows(2.8, 15.7, 2.55, 15.2, length = 0.06, col = "grey20")
text(2.5, 18.5, "Maximum r\nwith obs. CV = 0.2", pos = 4, cex = 0.9)
arrows(2.8, 19.3, 2.55, 19.7, length = 0.06, col = "grey20")
dev.off()
```
